Matching Bullets

### 2015-12-11 imaging-paper.Rnw
color scheme

\begin{figure}[hbtp]
\centering
\begin{subfigure}[t]{\linewidth}
\caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 35.\label{fig:smoothfac1}}{%
<<smoothfac1, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
br111.groove <- get_grooves(br111_bullet)
br111.loess <- fit_loess(br111_bullet, br111.groove)
get_peaks(br111.loess$data, smoothfactor = 35)$plot
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 5.\label{fig:smoothfac2}}{%
<<smoothfac2, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
get_peaks(br111.loess$data, smoothfactor = 5)$plot
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
    \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 65.\label{fig:smoothfac3}}{%
<<smoothfac3, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
get_peaks(br111.loess$data, smoothfactor = 65)$plot
@
    }
\end{subfigure}
\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 35, 5, and 65 respectively. Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 65 over-smoothes and causes the peaks/valleys to either disappear or be horizontally shifted from their position in the original loess residuals.}
\end{figure}

@@

By suggesting an automated algorithm that first identifies lands, then identifies peaks and valleys on this land, we reduce subjectivity and with it possible sources of bias. [T]`he concept of counting striations is subjective and based on experience, but it does tend to be consistent within a single examiner'~\cite{miller:1998}.
For a fair assessment of the adequacy of an algorithm we need transparency. Our matching  algorithm is open: the code is readily available in form of the R package x3rplus~\cite{x3prplus}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}. Obviously, this is only a first step, but in times of the Netflix prize~\cite{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle, we should not let the quality of the matching algorithm be the problem that stands in the way.

>>>>

\begin{figure}[hbtp]
%\centering
%\begin{subfigure}[t]{\linewidth}
%\caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 35.\label{fig:smoothfac1}}{%
<<smoothfac1, echo=FALSE, fig.width=7, fig.height=6, out.width='\\linewidth', warning=FALSE>>=
br111.groove <- get_grooves(br111_bullet)
br111.loess <- fit_loess(br111_bullet, br111.groove)

peaks1 <- get_peaks(br111.loess$data, smoothfactor = 15)
peaks1$lines$smoothfactor = 15
peaks1$plot$data$smoothfactor = 15
peaks2 <- get_peaks(br111.loess$data, smoothfactor = 5)
peaks2$lines$smoothfactor = 5
peaks2$plot$data$smoothfactor = 5
peaks3 <- get_peaks(br111.loess$data, smoothfactor = 35)
peaks3$lines$smoothfactor = 35
peaks3$plot$data$smoothfactor = 35

peaks <- rbind(peaks1$lines, peaks2$lines, peaks3$lines)
smooths <- rbind(peaks1$plot$data, peaks2$plot$data, 
                 peaks3$plot$data)

  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(smoothfactor~., labeller="label_both") +
    geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
              colour="grey50") +
    geom_line(aes(x=y, y=smoothed), data=smooths) +
  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) 
@
%'     }
%' \end{subfigure}    
%' \begin{subfigure}[t]{\linewidth}
%'     \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 5.\label{fig:smoothfac2}}{%
%' <<smoothfac2, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
%' get_peaks(br111.loess$data, smoothfactor = 5)$plot
%' @
%'     }
%' \end{subfigure}
%' \begin{subfigure}[t]{\linewidth}
%'     \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 65.\label{fig:smoothfac3}}{%
%' <<smoothfac3, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
%' get_peaks(br111.loess$data, smoothfactor = 65)$plot
%' @
%'     }
%' \end{subfigure}
\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 35, 5, and 65 respectively. Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 65 over-smoothes and causes the peaks/valleys to either disappear or be horizontally shifted from their position in the original loess residuals.}
\end{figure}

@@

By suggesting an automated algorithm that first identifies lands, then identifies peaks and valleys on this land, we reduce subjectivity and with it possible sources of bias. [T]`he concept of counting striations is subjective and based on experience, but it does tend to be consistent within a single examiner'~\cite{miller:1998}.

For a fair assessment of the adequacy of an algorithm we need transparency. Our matching  algorithm is open: the code is readily available in form of the R package x3rplus~\cite{x3prplus}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}. Obviously, this is only a first step, but in times of the Netflix prize~\cite{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle\footnote{kaggle home:\url{https://www.kaggle.com/solutions/competitions}}, we should not let the quality of the matching algorithm be the problem that stands in the way.
